% !TeX root = ../../tesi.tex

After acquisition and automatic labelling, the DEIMv2 detector is trained offline on a workstation with higher computational resources than the embedded devices used on site.\cite{huang_real-time_2026} This choice is necessary because the edge platforms are intended primarily for acquisition and online inference, whereas detector training requires longer optimization runs and larger memory budgets.
The automatically labelled images and the corresponding COCO-format annotations described in the previous section are used to build the detector datasets, which are then partitioned into training, validation, and test subsets with an 80/10/10 split.

Although the two industrial sites lead to different downstream tasks, the role of DEIMv2 is the same in both cases: the network must localize the product instances and provide reliable regions of interest for the subsequent processing stages. The model is therefore trained on two site-specific datasets, while preserving the same detector architecture and the same general training logic.\cite{labby_thesis_repo_2026} At Site~1, the detected regions are associated with aligned depth data and used for volumetric reconstruction. At Site~2, they are converted into loaf crops that feed the anomaly-detection branch. In this sense, DEIMv2 constitutes, together with the automatic labelling stage, the common detection layer of the overall framework.

During training, the optimization dynamics are monitored through TensorBoard, and the final checkpoint is selected according to the lowest validation loss observed during the training schedule. The resulting curves and the quantitative detection metrics obtained on the two datasets are discussed in Chapter~5.

Once the training stage is completed, the selected checkpoint is redeployed on the embedded target platform and becomes the detector used during online acquisition and inference. This offline-to-online transition allows the computationally demanding optimization stage to remain separated from the lightweight deployment stage, while preserving a consistent detection pipeline across the two monitoring sites.\cite{labby_thesis_repo_2026}
