% !TeX root = ../../tesi.tex

\subsubsection{Background Subtraction for Motion Triggering}
During the data-acquisition phase, the images are not recorded continuously. Instead, the acquisition framework monitors the scene through a configurable background-subtraction stage and stores a frame only when motion is detected inside the selected region of interest.\cite{opencv_background_subtraction_2026,labby_kinect_acquisition_2026} In this way, the saved dataset is concentrated around the actual passage of dough balls or bread loaves, while long intervals of static background are discarded.

The implementation supports the two OpenCV subtractors MOG2 and KNN, both wrapped inside a motion-detection interface that computes a foreground mask and converts it into a normalized motion score.\cite{opencv_background_subtraction_2026,labby_kinect_acquisition_2026} This score is then compared with configurable thresholds in order to decide whether the current frame is relevant enough to be stored. The main parameters of the acquisition logic, such as the subtractor type, history length, variance threshold, motion threshold, ROI geometry, image resolution, and initial warm-up offset, are collected in a YAML configuration file so that the system can be retuned without modifying the source code.\cite{labby_kinect_acquisition_2026}

The same logic is reused across the two sites, but the sensing modality depends on the available hardware. In the Kinect branch, ROI-based triggering can be performed on the transformed depth image, which is especially convenient when the monitored product moves on top of a stable support surface. In RGB-only branches, such as the Raspberry Pi acquisition setup used at Site~2, the foreground mask is instead computed directly from the RGB stream.\cite{labby_kinect_acquisition_2026} The framework also skips an initial number of frames in order to stabilize the background model before the actual recording starts, thereby reducing false triggers caused by transient initialization artefacts.

When the motion score exceeds the configured threshold within the selected ROI, the current frame is written to disk together with the modality-specific data associated with the active sensor. The codebase also supports the automatic creation of timestamped folders when motion starts, so that separate acquisition sessions can be organized without manual intervention.\cite{labby_kinect_acquisition_2026} This motion-triggered acquisition strategy proved particularly suitable for long autonomous campaigns, because it reduces redundant storage while preserving the samples that are most informative for the subsequent labelling and training stages.
