% !TeX root = ../../tesi.tex

\chapter{Conclusions and Future Work}
This thesis has presented a multimodal quality-inspection framework for the bread-production process. The proposed system combines motion-triggered acquisition, automatic labelling, DEIMv2-based object detection, volumetric reconstruction of dough balls from RGB-D data, and anomaly detection for baked products. The implementation was developed in close interaction with the operators of the SMACT \cite{smact_official_2026} Competence Center, so that the final solution could remain aligned with the practical constraints of the industrial environment as well as with the actual monitoring needs of the production line, while requiring no modifications to the existing production setup and operating without the need for additional infrastructure.

The results obtained in the two monitored sites highlight the effectiveness of the proposed framework. At Site~1, the volumetric analysis showed that the rotation-based methods clearly outperform the simpler depth-gap baseline, with the rotated convex-hull estimator providing the best agreement with the expected reference interval. The experiments also showed that, although SAM can provide very accurate segmentations, a lighter color-clustering plus background-subtraction strategy represents a more suitable and robust compromise for the intended workflow, since it preserves good reconstruction quality while remaining computationally more efficient.
At Site~2, the anomaly-detection branch achieved a near-perfect image-level AUROC of 0.999958 under the adopted evaluation protocol, while also showing consistently low anomaly scores on normal bread samples and meaningful responses on external anomalous examples, with anomaly maps that correctly localized the most relevant irregular regions.

Overall, the quantitative and qualitative results demonstrate that the framework is capable of supporting quality monitoring in a real industrial environment, across both the dough-preparation and baked-product stages.

Future work may focus on improving both the accuracy and the industrial robustness of the proposed framework. For Site~1, a more reliable reference for volume evaluation could be obtained through direct volumetric measurements on representative dough samples, thus reducing the uncertainty associated with the current indirect estimate based on nominal weight and density. For Site~2, the collection of a broader and more representative set of defective bread samples would make it possible to validate the anomaly-detection branch more extensively and to better characterize its behavior under realistic fault conditions. Additional developments may also include even tighter real-time integration with the production process, so as to provide immediate feedback to operators.
